---
title: Supplementary Notes for *R for Data Science*
author: Jeffrey Arnold
editor_options: 
  chunk_output_type: console
---

# Data Visualization

# Workflow: basics

# Data Transformation

# Exploratory Data Analysis

# Workflow: Projects

# Variable names in R

# Tibbles

- What can variables be named in R?
- What are non-syntactic variables?

# Data import

- What are csv files?
- How does a csv reader determine the types of variables?
- Read Stata and SPSS files

## Notes

Also see 

# Tidy data

Objectives

- Describe what "tidy data" frame is and why it is useful
- Describe when non-tidy data formats may be useful
- Convert data from long to wide formats using `spread()` and `gather()`
- Combine and split variables using `unite()` and `gather()`

## Notes

- This is an especially import chapter. Read it carefully.

- What is the difference between data for storage and for analysis? Why might analytical data not be in a tidy format?
- COW Data - What format is it?

## Summarize multiple variables

Reshaping a data frame to a long format with columns for variables and values is a convenient format to calculate summary statistics for multiple variables.
```{r results='asis'}
library("tidyverse")
library("gapminder")

gapminder_summary <-
  gapminder %>%
  select(lifeExp, pop, gdpPercap) %>%
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(mean = mean(value), sd = sd(value))
gapminder_summary
```

If necessary, the data can be reshaped so each (variable, statistic) combination is a column:
```{r}
gapminder_summary %>%
  gather(stat, value, -variable) %>%
  # create column with columns variable, stat, value
  unite(variable_stat, variable, stat) %>%
  # combine variable and stats into a single column
  spread(variable_stat, value)
  # spread into a single column
```

An alternative method to summarize multiple variables is to use the `summarise_at` function:
```{r}
summarise_at(gapminder, vars(gdpPercap, lifeExp, pop), 
             funs(mean, sd))
```

## Faceted Variables

If multiple variables need to be plotted, an alternative to creating multiple plots is to create a long data frame with a variable column and then facet by variable. 
For example, to plot the distribution of several variables in the same plot:
```{r}
library("tidyverse")
library("gapminder")

gapminder_long <-
  gapminder %>%
  select(lifeExp, pop, gdpPercap) %>%
  gather(variable, value)

ggplot(gapminder_long, aes(x = value)) +
  geom_density() +
  facet_wrap(~ variable, scales = "free", ncol = 1)
  scale_x_log10()


```


# Character vectors

- What is character encoding?
- What are escape sequences?
- What is a regular expression?

## Regular Expressions

Know what they are ... 

- match patterns
- find and replace patterns of strings 
- extract parts from string


Examples of uses:

- match filenames
- filter character vectors
- select columns using `match()`
- rename column names using `rename_all()`

```{r}
text <- readLines(system.file("extdata", "federalist", "fp01.txt", 
                              package = "qss")) %>%
  str_c(collapse = " ") %>%
  str_match_all("[A-Za-z]+") 

text <- tibble(text = text[[1]][, 1]) %>%
  mutate(text = str_to_lower(text)) %>%
  count(text) %>%
  arrange(desc(n))
```


```{r}

```


## Encoding

Character encoding can occur when reading data from csv and text files.

Character encoding problems will occur at some point. It is becoming less of an issue due to a standardization to UTF-8, but it still occurs frequently.
The most likely encoding that you will encounter (if working with European texts) is Windows-1252 (CP-1252) or ISO-8859-1.

## Escape sequences

Certain characters preceeded by a backslash are treated specially. The most common are: `\n` (new-line), `\t` (tab).

Use a double-backslash `\\` for backslash in character vectors.

One common error is using a backslash in Windows file paths, which results in an error:
```{r error=TRUE}
"path\to\something"
```
Escaping the backslash will work, but won't work on MacOS or Linux.
It is better to either always use forward-slashes or `file.path` when writing paths in R.
```{r}
file.path("path", "so", "something")
file.path("path/to/something")
```

### Exercises

1. Programatically create a variable name
2. Rename a variable
3. Extract parts from U.S. Congressperson titles, e.g. "Ryan (R-WI)"
4. Find all .csv files in a list of names.


# Dates and Times


- Convert date vectors to and from strings
- How are dates stored?

Date-time variables are almost always a pain to work with.
Common issues include the irregularity of months and time-zones.

Some alternative date and time packages

- What is the format of dates?
- Convert to and from dates
- Time-zones and such

# Factors

- Difference between character vector and factor?
- When to use factors?

    - You need to store metadata, e.g. levels
    - Control over the order of categories for plotting in ggplot
    - Models, e.g. `lm`

- http://stat545.com/block029_factors.html

# Lists and Vectors

- Lists vs. atomic vectors
- Matrices vs. data frames
- Issues with floating point numbers? Why are they called "double" in tidyverse functions? 

When to use lists? If you have several variables 



