---
title: How to avoid copying and pasting code by using functions, for loops and functionals
---

One of the most important principles in writing software is "don't repeat yourself (DRY)".
This module covers using functions, for-loops, and functionals to avoid copying and pasting code.

In addition to this, see the the *R for Data Science* chapters [Functions](http://r4ds.had.co.nz/functions.html) a [Iteration](http://r4ds.had.co.nz/iteration.html) (for-loops and functions) for more detailed discussions of these concepts.


## Setup {-} 

Consider the example from QSS Exercise 1. Review the explanations and instructions for that question.
```{r}
# a QSS Exercise from Chapter 1.
library("tidyverse")
library("qss")

data("Kenya", package = "qss")
data("Sweden", package = "qss")
data("World", package = "qss")
```

## Replacing Copy and Pasting Code with Functions

On question asks you to calculate the CDR (crude death rate)
```{r}
# Let's write some code that works for Kenya
Kenya %>%
  group_by(period) %>%
  summarize(CDR_data = sum(deaths) / sum(py.men + py.women))

```

Awesome! This code does what I want.

But, it only works for Kenya, and I'd need to copy and paste it to work on the
`Sweden` and `World` data frames.
```{r}
Sweden %>%
  group_by(period) %>%
  summarize(CDR_data = sum(deaths) / sum(py.men + py.women))

World %>%
  group_by(period) %>%
  summarize(CDR_data = sum(deaths) / sum(py.men + py.women))
```

R4DS notes that I should write a function anytime I'm doing the same thing [3 or more times](http://r4ds.had.co.nz/functions.html#when-should-you-write-a-function).

So to generalize the previous code, consider what changes when copy and pasting
this data.
The only thing that changes is the specification of the input data frame: `Kenya`, 
`Sweden`, or `World`. I could rewrite this code by creating a new variable `x`
which I can assign to the input data frame I want to use.

```{r}
x <- Kenya
# x <- Sweden
# x <- World

x %>%
  group_by(period) %>%
  summarize(CDR_data = sum(deaths) / sum(py.men + py.women))
```

This helps clarify the programming task, but isn't yet that useful because we would
still have to copy and paste that code to run it in a script.

However, now that we've written the code such that the input is a variable `x` and an expression
of code using that variable `x`.
We can take that previous code and encapsulate it in a function.
```{r}
CDR <- function(x) {
  x %>%
    group_by(period) %>%
    summarize(CDR_data = sum(deaths) / sum(py.men + py.women))
}
```

A function consists of there parts: `name`, `arguments` (inputs), and `body`.

1. Functions are objects. You can assign them to variable names. `name` is the name you assign the function to. In this case, `CDR`.
2. The `arguments` are names that are given to input values. 
3. The `body` is the code that the function executes.

Now we can run the `CDR` on Kenya, Sweden, and World:
```{r}
CDR(Kenya)
```
```{r}
CDR(Sweden)
```
```{r}
CDR(World)
```

## Repeating Inputs with For Loops

In the previous section we still had to run the function `CDR` three times, once for each
country. 
While three countries is not terrible, this would be quite tedious if we had hundreds of countries.

In this section, we'll reduce that redundancy using a for loop.

### Lists

The first step is to put those data frames in a list.
```{r}
countries <- list(KEN = Kenya, SWE = Sweden, WLD = World)
```
The names of the elements of the list are arbitrary, but were purposefully chosen 
to be different than the names of the data frames to emphasize that they are arbitrary.

The list `countries` has `r length(countries)` elements,
```{r}
length(countries)
```
with names
```{r}
names(countries)
```

To get an element from the list, use `[[` or `$` and reference the elements by name or index number:
```{r}
summary(countries$KEN)
summary(countries[["SWE"]])
summary(countries[[3]])
```

The for loop lets us run 
```{r}
for (x in countries) {
  print(glimpse(x))
  print(CDR(x))
}
```

The for loop runs three times:

1. Set `x = countries[[1]]`, and run `CDR(x)`
2. Set `x = countries[[2]]`, and run `CDR(x)`
3. Set `x = countries[[3]]`, and run `CDR(x)`

However, while we've run `CDR` on each country we haven't saved the results to use anywhere.
We'll make three changes to our code:

1. Define an empty list `cdr_results` of the same length as `countries` to store the results
2. Loop over the names of the countries instead of the values so we can name the elements of the results.
3. Within the loop save the result to an element in `cdr_results`

First, create an empty vector the same length as countries.
```{r}
cdr_results <- vector("list", length = length(countries))
```
```{r}
cdr_results
```

Then,  loop over each country, running `CDR` on that country's data frame and saving 
the results to an element of `countries`:
```{r}
for (i in names(countries)) {
  cdr_results[[i]] <- CDR(countries[[i]])
}
```

Now, `cdr_results` contains the results of running `CDR` on each of those data frames:
```{r}
cdr_results[["KEN"]]
```
```{r}
cdr_results[["SWE"]]
```
```{r}
cdr_results[["WLD"]]
```


## Replacing for loops with map functions

One annoyance and possible place to make bugs in for loops is that we need to define
a vector to store the results.
This requires that we create the vector with the correct length, and remember to 
update the same vector.

The map functions in the **purr** package are functions that apply a function to each
element of a vector.

The function `map(.x, .f)` applies the function `.f` to each element of the vector `.x`.
The result is
```
list(.f(.x[[1]]), .f(.x[[2]]), ..., .f(.x[[length(.x)]]))
```

Using the `map` function  we can replace the previous for loop with a single line of code:
```{r}
map(countries, CDR)
```
This applied the function `CDR` to each data frame in the `countries` list and returned another list
with all the results. Unlike using a for loop we did not need to first create an empty vector.
