---
title: "Git and GitHub"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup,include=FALSE}
library("DiagrammeR")
```

## Cloning, Pull Requests, Forking, Pulling, and Pushing

<img alt="Relationship between git and GitHub" src="/2018/img/github-diagram.svg">

See the *Happy Git and GitHub for the useR* chapter [Fork a repo](http://happygitwithr.com/fork.html).

## Homework Workflow

```{r echo=FALSE}
mermaid("homework-flow.mmd")
```

## Project workflow

For the project assignments, each student has a repository named
`UW-POLS501/:username` where `:username` is the GitHub username of the student.
This does not need to be forked.
The student works directly on that repository, and submits assignments by opening an issue on `UW-POLS501/:username/issue`.

```{r echo=FALSE}
mermaid("project-workflow.mmd")
```



## References

- Jenny Bryan, [Happy Git and GitHub for the useR](http://happygitwithr.com/)
-   DataCamp course [Version Control with Git](http://swcarpentry.github.io/git-novice/). This course requires shell knowledge, but the following two modules are useful.
    - [Automated Version Control](http://swcarpentry.github.io/git-novice/01-basics/)
    - [Using Git from RStudio](http://swcarpentry.github.io/git-novice/14-supplemental-rstudio/)
- [Simple guide to forks in GitHub and Git](http://www.dataschool.io/simple-guide-to-forks-in-github-and-git/)
- [GitHub Help](https://help.github.com/)
- [Git and GitHub videos for beginners](http://www.dataschool.io/git-and-github-videos-for-beginners/) from dataschool introduces git and GitHub via many videos.
